Exploiting the Social Graph: Increasing Engagement in Collaborative Interactive Evolution

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Adding future lines of work
Which link this to other lines of research. You can add more, of
course.

Besides, this framework is a proposal to view IEC as human centered computing [cite],
in which the context, environment, interfaces, preferences, accesibility, human relations, cognitive
limitations, culture, creativity and other human beings aspects are an integral part of the system. 
This view requires a multidisciplinary approach, as well as multiple knowledge representation formalisms.
As a first step toward, the traditional data model is extended with a graph, in order to have 
yet another view to exploit. Following a description of the data model is presented.

@@

A graph is composed of two fundamental entities: nodes and relationships,
both having attributes. Nodes are abstractions of the objects in the real 
world system, and relationships describe their interactions. Types describe
the attributes describing an entity.

@@

\item {\bf Collection} This type of node represents the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
belonging to this collection can be shared with others. Collections can contain
phenotypes and a single phenotype could be shared by many
collections. % This is why they are not attributes - Mario
% You can simply have an attribute that is a list of the collections
% it belongs to - JJ
% Right, but:
%  1) The relationship could have additional atributes, (when it was stored, order, etc.)
%  2) Then it could be difficult to query with Cypher (I think) - Mario
\end{itemize}

@@

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
a phenotype in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
a phenotype multiple times, so the date of the interactions is also of
interest.
%this is a bipartite graph them. - JJ
% Do you mean like a figure?  - Mario
% No, it's a kind of graph - JJ
% I read graph them :)  but is:  graph then , now I get it.

\item {\bf Knows} The relation connects two users that know each
other. 
% This relation is "unique" meaning that it cannot be repeated
% between a pair of nodes.  % Not too clear what you mean by that - JJ
% I mean is not a multigraph or you can not have two equal relationships between nodes. - Mario
 % Watching the graph structure of Neo4J, I can now see that is actually a directed graph,
 % There is even cases where knows relationship only goes in one direction, this could be because
 % a user does not publishes their friends or one only follows another but are not friends of each other.  
% Deleting the restriction, I think is not important and can be seen
% as a graph.
% It's a directed, bipartite graph then - JJ
% Ok, I will mention the types of graphs. -Mario

@@

Is important to notice that this model is still a partial view of the IEC system,
and other models can could be used in practice when implementing the
system. % It is also a bipartite graph, which could be converted to a
% single-party graph with only one type of node. 
% I don't get this comment, yes some relations are bipartite, but If we see the whole
% graph I think is a multi-graph.   - Mario 
% I will check the  single-party graph conversion you suggest, I don't know the concept. - Mario

% I don't think "collections" should be nodes too. They are simply
% attributes for an individual - JJ
% They are nodes, I added the reason above. - Mario
For instance a relational database was used for the implementation of a 
basic user profile and authentication module. Also the active population 
was kept in a key-value store in memory for performance reasons.  
Nevertheless, having a multi-model implementation can give important 
insights of the behavior of the system. This architecture was validated
by a case study, requiring the interaction of all of the components.        
 % So what we have done here is this and that because... - JJ
 % Changed the paragraph. - Mario 
% refinement
\section{Case Study}
\label{sec:experiments}

@@

Only the ``Start a session'' and ``Explore collections of other friends'' actions 
are not reflected in the graph model since they are stored instead in the PostgreSQL relational
database system. This is because user sessions and authentication, as well as basic
web requests, are handled directly by the django web framework \cite{garcia2013evospace}.  
     % because... - JJ
     % done - Mario

Two scores are used to determine the weight of a user's preference:
\begin{itemize}
\item Experience. This value depends on the score and is a value 
between 0 and 100. It is assumed for this case, that once a user
reaches a 100 actions, it has enough experience on using the application.   
% More details needed? Is done by a Fuzzy Inference System but I think
% the description would take to much space.
% Better a small description than nothing. If you need space, some
% sections above are maybe a bit too verbose - JJ
% I checked and this is not a fuzzy value. Explained any way. - Mario

\item Participation. This value is simply the degree of the user node 
in the graph. This measure considers the number of users he knows,
collections and phenotypes rated.    
\end{itemize}

@@

\subsubsection{ Non Graph Gamification deployment}
The total number of volunteers for this experiment was very close to those obtained in the previous
experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
with a total of 2596 relationships. These numbers are close to those obtained in the base experiment,
with a marginal increase in participation. The increased participation also helped in the creation of
more phenotypes for a total of 648. The total number of evaluated
phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} in red.

@@

When monitoring the experiments by querying the graph database,  the Cypher query language 
provided interesting views on the data that it would have been difficult to replicate with a relational database.
For instance, to compare the number of volunteers in deployments
B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively)  a simple query
provided a view of the social network of volunteers. 
%Please read the whole sentence and maybe rewrite - JJ
% Done - Mario

@@

>>>>

Besides, the methodology presented in this paper for interactive
evolutionary computation is a proposal to frame it as human centered computing \cite{sebe2010human},
in which the context, environment, interfaces, preferences, accesibility, human relations, cognitive
limitations, culture, creativity and other human aspects are an integral part of the system. 
This view requires a multidisciplinary approach, as well as multiple
knowledge representation formalisms. 
As a first step forward, the traditional data model is extended with a graph, in order to have 
yet another view to exploit. In the following section a description of
the data model is presented.

@@

A graph is composed of two fundamental entities: nodes and edges,
both having attributes. Nodes are abstractions of the objects in the real 
world system, and relationships describe their interactions. Types describe
the attributes describing an entity, and edges express relationships
between nodes.

@@

\item {\bf Collection} This type of node represents the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
belonging to this collection can be shared with others. Collections can contain
phenotypes and a single phenotype could be shared by many
collections. % This is why they are not attributes - Mario
% You can simply have an attribute that is a list of the collections
% it belongs to - JJ
% Right, but:
%  1) The relationship could have additional atributes, (when it was stored, order, etc.)
%  2) Then it could be difficult to query with Cypher (I think) -
%  Mario
% Then it's more of an implementation detail, but anyway, point taken
% - JJ
\end{itemize}

@@

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
a phenotype in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
a phenotype multiple times, so the date of the interactions is also of
interest.

\item {\bf Knows} The relation connects two users that know each
other.

@@

Is important to notice that this model is still a partial view of the IEC system,
and other models can could be used in practice when implementing the
system. % It is also a bipartite graph, which could be converted to a
% single-party graph with only one type of node. 
% I don't get this comment, yes some relations are bipartite, but If we see the whole
% graph I think is a multi-graph.   - Mario 
% I will check the  single-party graph conversion you suggest, I don't
% know the concept. - Mario
% Does not matter much... it converts a bipartite graph to another
% graph with a single kind of nodes, phenotypes or users, for instance
% - JJ

% I don't think "collections" should be nodes too. They are simply
% attributes for an individual - JJ
% They are nodes, I added the reason above. - Mario
For instance a relational database was used for the implementation of a 
basic user profile and authentication module. Also the active population 
was kept in a key-value store in memory for performance reasons.  
Nevertheless, having a multi-model implementation can give important 
insights of the behavior of the system. This architecture was validated
by a case study, requiring the interaction of all of the components.

\section{Case Study}
\label{sec:experiments}

@@

Only the ``Start a session'' and ``Explore collections of other friends'' actions 
are not reflected in the graph model since they are stored instead in the PostgreSQL relational
database system. This is because user sessions and authentication, as well as basic
web requests, are handled directly by the Django web framework \cite{garcia2013evospace}.

Two scores are used to determine the weight of a user's preference:
\begin{itemize}
\item {\em Experience}: This value depends on the score and is a value 
between 0 and 100. It is assumed for this case, that once a user
reaches a 100 actions, it has enough experience on using the application.

\item {\em Participation}: This value is simply the degree of the user node 
in the graph. This measure considers the number of users he knows,
collections and phenotypes rated.    
\end{itemize}

@@

\subsubsection{ Non Graph Gamification deployment}
The total number of volunteers for this experiment was very close to
the one obtained in the previous
experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
with a total of 2596 relationships. These numbers are close to those obtained in the base experiment,
with a marginal increase in participation. The increased participation also helped in the creation of
more phenotypes for a total of 648. The total number of evaluated
phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} in red.

@@

When monitoring the experiments by querying the graph database,  the Cypher query language 
provided interesting views on the data that it would have been
difficult to replicate with a relational database. 
For instance, to compare the number of volunteers in deployments
B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively)  a simple query
provided a view of the social network of volunteers.

@@

One of the interesting future lines of work would be to look a bit
more closely at the behavior of users rating the system. These initial
experiments hint at a possible power law, which might indicate that
the IEC system could be self-organizing, a process that would allow it
to reach a critical state, as has been found in software repositories,
for instance \cite{Merelo2016:repomining}. The dynamics of this kind
of system are fundamentally different, and our future research will
include exploring these aspects of the system.
